---
title: "Bike Share and Weather Data in Washington DC"
author: "STAT 405 - Vivian Zheng, Tyler Braito, Calvin Aberg, Mattia Saladini, and Ndidi Nwosu"
date: "November 29, 2022"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Loading data:
#setwd("~/Documents/Sem 1/STAT 405/City Bike Data")
library(ggplot2)
library(dplyr)
library(ggthemes)
library(RSQLite)
library(stringr)
library(lubridate)
library(ggeffects)
library(dotwhisker)
library(ggeffects)
library(margins)
library(geosphere)
library(grid)
library(tidyverse)
# install.packages('sf')

# Import bikeshare and weather csv as SQL
dcon <- dbConnect(SQLite(), dbname = "mydba.sqlite")
weatherTable <- read.csv(paste0("C:/Users/Ndidi/Desktop/2022-23/fall/STAT 405/data project/stat405-project/Washington,DC,USA weather 2022-08-01 to 2022-08-31.csv"))
dbWriteTable(conn = dcon, name = "weatherTable", 
             weatherTable, append = TRUE, row.names = FALSE)
dbListTables(dcon)
bike_dta <- read.csv("C:/Users/Ndidi/Desktop/2022-23/fall/STAT 405/202208-capitalbikeshare-tripdata/202208-capitalbikeshare-tripdata.csv", header = TRUE, sep = ",")
dbWriteTable(conn = dcon, name = "bike_dta", 
             bike_dta, append = TRUE, row.names = FALSE)
dbListTables(dcon)
```
## Our Data

* **Primary Dataset**: City Bike Data in Washington D.C, August 2022
* **Secondary Dataset**: Weather Data in Washington D.C., August 2022

## Our Research
* **Research Questions:**
  + Is there a pattern in what bike locations are most popular, or a correlation between the type of bike share member, the duration of the ride or the type of bike chosen?
  + How does the weather affect the number of people using city bikes?

* **Implications of Research:**
  + Beneficial to city bike share company to see how to maximize the amount of customers
  + Beneficial to users to see trends in business based off of their own start stations

## Scatterplot of Ride Duration by Date

```{r, echo=FALSE, warning=FALSE}
# library(ggplot2)
# install.packages("tidyverse")
start_time <- bike_dta$started_at
end_time <- bike_dta$ended_at
bike_dta$start_time <- as.POSIXct(start_time,tz="EST")
bike_dta$end_time <- as.POSIXct(end_time,tz="EST")
bike_dta$duration <- as.numeric((bike_dta$end_time - bike_dta$start_time) / 60)

# Identify top station (Lincoln Memorial)
# tab <- table(bike_dta$start_station_name)
# sort(tab)
linc_mem <- subset(bike_dta, bike_dta$start_station_name == "Lincoln Memorial" &
                     bike_dta$end_station_name != "Lincoln Memorial")
linc_mem <- na.omit(linc_mem)

ggplot(data = linc_mem, aes(x = start_time, y = duration)) +
  geom_point(color = "skyblue") +
  ggtitle("Duration by Date for Lincoln Memorial Station, August 2022") +
  xlab("Date") + ylab("Rental Duration (min)")
```


## Daily Usage Rank of the Top 10 Most Common Starting Stations


```{r, echo=FALSE, warning=FALSE}
# library(ggplot2)
# library(dplyr)
bike_dta$date <- format(bike_dta$start_time, '%Y-%m-%d')
bike_data_sub <- subset(bike_dta, select = c("start_station_name", "end_station_name", "date"))
bike_data_sub <- bike_data_sub %>% 
  add_count(start_station_name, name = 'start_occurance')
bike_data_sub <- bike_data_sub %>%
  add_count(end_station_name, name = 'end_occurance')
# View(bike_data_sub)

#Only use 10 most common stations
bike_data_sub <- subset(bike_data_sub, end_station_name %in% names(sort(table(bike_data_sub$start_station_name),decreasing=TRUE)[2:11]))
bike_data_sub <- subset(bike_data_sub, start_station_name %in% bike_data_sub$end_station_name)

length(unique(bike_data_sub$start_station_name))
length(unique(bike_data_sub$end_station_name))


bike_sub2 <- bike_data_sub %>%
  group_by(date) %>%
  add_count(start_station_name, name = 'occurance_by_day') %>%
  ungroup()
  
bike_sub2 <- bike_sub2 %>% distinct(start_station_name, date, .keep_all=TRUE)

bike_sub3 <- bike_sub2 %>%
  group_by(date) %>%
  arrange(date, desc(occurance_by_day), start_station_name) %>%
  mutate(rank=row_number())%>%
  ungroup()

bike_sub3$date <- as.Date(bike_sub3$date)


ggplot(bike_sub3, aes(x=date, y = rank, group = start_station_name))+
  geom_line(aes(color = start_station_name), size = 2, alpha = .75)+
  geom_point(aes(color = start_station_name), size = 4, alpha = .75)+
  scale_y_reverse(breaks = 1:10)+
  scale_color_brewer(palette = "Paired", name = "Station Name")+
  scale_x_date(date_breaks = "6 days")+
  theme(legend.position = "bottom") +
  guides(color = guide_legend(nrow = 4)) +
  labs(title = "Daily Usage Rank of the Top 10 Most Common Starting Stations", x = "Date", y = "Rank")
```


## Density Plot of Riding Distance by Bike Type

```{r, echo=FALSE, output=TRUE, warning=FALSE}
#Calculating duration:
# start_time <- as.POSIXct(bike_dta$started_at,tz="EST")
# end_time <- as.POSIXct(bike_dta$ended_at,tz="EST")
bike_dta$duration_sec <- bike_dta$end_time - bike_dta$start_time
bike_dta$duration_min <- bike_dta$duration_sec/60



#Densities:
bike_dta <- na.omit(bike_dta)
bike_dta$distance <-sqrt((69*(bike_dta$end_lat - bike_dta$start_lat))**2 + 55*((bike_dta$end_lat - bike_dta$start_lat))**2)
bike_dta2 <- bike_dta[bike_dta$distance<=3, ]

ggplot(bike_dta2, aes(x=distance)) + geom_density(alpha=.2, fill="#FF6666") + geom_vline(aes(xintercept=mean(distance)), color="blue", linetype="dashed", size=1) + labs(x="Distance (miles)", y = "Density") + facet_grid(. ~ rideable_type)
```

## Boxplots of Bike Types and Rental Durations


```{r, echo = FALSE, output = TRUE, warning=FALSE}
#Calculating duration:
# start_time <- as.POSIXct(bike_dta$started_at,tz="EST")
# end_time <- as.POSIXct(bike_dta$ended_at,tz="EST")
bike_dta$duration_sec <- bike_dta$end_time - bike_dta$start_time
bike_dta$duration_min <- bike_dta$duration_sec/60

#Boxplots:
par(mfrow = c(1, 3))
boxplot(subset(bike_dta$duration_min, bike_dta$duration_min<=60 & bike_dta$rideable_type == "classic_bike"), main="Classic Bikes Boxplot", 
        ylab="Duration (min)", col="red")
boxplot(subset(bike_dta$duration_min, bike_dta$duration_min<=60 & bike_dta$rideable_type == "electric_bike"), main="Electric Bikes Boxplot", col="green")
boxplot(subset(bike_dta$duration_min, bike_dta$duration_min<=60 & bike_dta$rideable_type == "docked_bike"), main="Docked Bikes Boxplot", col="blue")

```

## Most Popular Stations by Start and End Location

```{r, echo = FALSE, output = TRUE}
counts_start <- head(sort(table(bike_dta$start_station_name), decreasing = TRUE), 6)[-1]

counts_end <- head(sort(table(bike_dta$end_station_name), decreasing = TRUE), 6)[-1]

stations <- rbind(counts_start,counts_end)

par(mfrow = c(1, 1))
barplot(stations, col = c('darkblue', 'red'),
        names.arg = c("Lincoln",'Jefferson', "T St.", '15th', 'Union St'),
        main = "Most Popular Stations by Start and End Location",
        legend = c('End Location', 'Start Location'), 
        args.legend = list(x = "topright",
                           inset = c(0, -0.1)),
        xlab = 'Stations',
        ylab = "Count of Riders")
```


## Histogram of Usage Frequency by Days of the Week and Date

```{r, echo = FALSE, output = TRUE}
par(mfrow = c(1, 2))
# bike_dta$started_at <- as.POSIXct(bike_dta$started_at, tz = "EST")
# bike_dta$ended_at <-as.POSIXct(bike_dta$ended_at, tz = "EST")
# bike_dta$duration <-  as.numeric(bike_dta$end_time - bike_dta$start_time)
bike_dta$weekday <-as.Date(bike_dta$start_time)
bike_dta$weekday <- format(bike_dta$weekday, "%a")
# bike_dta$weekdayn <- as.numeric(format(bike_dta$start_time, "%w"))
# hist(bike_dta$weekdayn, breaks = -.5+0:7, labels = unique(bike_dta$weekday[order(bike_dta$weekdayn)]), xlab = "", xaxt = "n",  main ="Ride Frequency Per Day of Week", col = "deepskyblue")
barplot(table(bike_dta$weekday), main = "Ride Frequency Per Day of Week", xlab = "Day of the Week", col = "deepskyblue")
date_formatted <- as.Date(bike_dta$start_time, format = "%Y-%m-%d")
ride_start_days <- format(date_formatted, "%d")
barplot(table(ride_start_days), main = "Rides By Day in August", xlab = "Day", ylab = "Number of Rides", col = "yellow")

```



## Bike Rides vs Average Temperature of That Day

```{r, echo = FALSE, output = TRUE}
# weather <- read.csv("C:/Users/Ndidi/Desktop/2022-23/fall/STAT 405/data project/stat405-project/Washington,DC,USA weather 2022-08-01 to 2022-08-31.csv")
# Create query for average based on condition
# res <- dbSendQuery(conn = dcon, "
#               SELECT w.datetime, w.temp, b.started_at
#               FROM bikes b, weather w;")
# 
# # store results in df
# rides <- dbFetch(res, -1)
# dbClearResult(res)

rides_per_day <- data.frame(weatherTable$datetime, weatherTable$temp, table(ride_start_days))

ggplot(rides_per_day)  + 
  geom_bar(aes(x=ride_start_days, y=Freq),stat="identity", fill="lightgreen",color="darkgreen")+
  geom_line(aes(x=ride_start_days, y=weatherTable.temp *150, group = 1),stat="identity",color="red",size=2)+
  labs(title= "Bike Rides vs Temperature (in Fahrenheit)",
       x="day",y="Number of Bike Rides")+
  scale_y_continuous(sec.axis=sec_axis(
    ~./ 150,name="Temperature"))

```


## Concentration of Trip Starts

```{r, echo = FALSE, output = TRUE, warning=FALSE, message=FALSE}
# Create query for average based on condition
res <- dbSendQuery(conn = dcon, "
              SELECT start_station_name, end_station_name, COUNT(*) as num_trips
              FROM bike_dta b
              GROUP BY start_station_name, end_station_name
              ORDER BY (-1 * num_trips) 
              LIMIT 10;
                   ")

# store results in df
station_travel <- dbFetch(res, -1)
dbClearResult(res)
# station_travel

ggplot(data = station_travel[-1,], aes(x = start_station_name, y = num_trips)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  ggtitle("Concentration of Trip Starts") +
  labs(x = "Start Station Name", y = "Number of Trips") +
  theme_minimal()
```



## Effects of Afternoon Rain on Bike Usage

```{r, echo = FALSE, output = TRUE, warning=FALSE}
## Join two dataframes based on start date
res <- dbSendQuery(conn = dcon, "
SELECT *
FROM (SELECT *, date(started_at) as date FROM bike_dta) as a, weatherTable
WHERE a.date = weatherTable.datetime
ORDER BY a.date")
mydf <- dbFetch(res,-1)
dbClearResult(res)

## Search Description for presence of After Noon Rain
mydf$after_noon_rain <- str_detect(mydf$description, "afternoon rain")
mydf$after_noon_rain <-  factor(mydf$after_noon_rain, labels = c("No Afternoon Rain", "Afternoon Rain"))
 
#Extract Hour From Time Stamp 
mydf$hour <- hour(mydf$started_at)

#Plot it
ggplot(data = mydf, aes(x = hour))+
  geom_histogram(stat = "count", fill = "deepskyblue3")+
  facet_wrap(~after_noon_rain, scales = "free")+
  labs(title = "Effects of Afternoon Rain on Bike Usage", y = "Trip Count", x = "Hour of Day", 
       facet = c("No Rain", "Rain"))+
  theme_bw()
```


## Effects of Weather Description on Bike Rides

```{r, echo = FALSE, output = TRUE, warning=FALSE, message=FALSE}
weatherTable$cloudy <- as.factor(as.numeric(str_detect(weatherTable$description, "cloud")))
weatherTable$clear <- as.factor(as.numeric(str_detect(weatherTable$description, "Clear")))
weatherTable$rain <- as.factor(as.numeric(str_detect(weatherTable$description, "rain")))

bike_dta$date <- str_sub(bike_dta$started_at, 1, 10)

# aggregating by rides per day
by_date <- bike_dta %>%
  group_by(date) %>%
  summarise(total_rides = n()) 
weatherTable$date <- as.Date(weatherTable$datetime)
by_date$date <- by_date$date
dates <- by_date$date
by_date$date <- as.POSIXct(dates,tz=Sys.timezone())
# joining weather data with bike data
ride_weather <- inner_join(by_date, weatherTable)

mod <- lm(total_rides ~ rain+clear+cloudy, data = ride_weather)
# summary(mod)
# creating a substantive effect plot
dwplot(mod, vline = geom_vline(xintercept = 0, colour = "grey60", linetype = 2))%>% 
  relabel_predictors(c(rain1= "Rain", clear1 = "Clear",cloudy1 = "Cloudy")) +
  ggtitle("Effects of Weather Description Factors on Rides")

```


## Modeling

```{r, include=FALSE, warning=FALSE, message=FALSE}
# bikes
# data manipulation
bike_dta$date <- substr(bike_dta$started_at, 1, 11)

mod_data <- bike_dta %>%
  group_by(date) %>%
  summarize(num_bikers = n())

# mod_data$date <- as.Date(mod_data$date)
mod_data$date <- str_trim(mod_data$date)

mod_data <- mod_data %>% inner_join(weatherTable, by = c('date' = "datetime"))
```

```{r, include=FALSE, warning=FALSE, message=FALSE}
# linear modeling
mod_1 = lm(num_bikers~tempmax+tempmin+temp+feelslike+dew+humidity+precip+windgust+windspeed+winddir+visibility+solarradiation+severerisk+uvindex, data = mod_data)
# summary(mod_1)
```


```{r, include=FALSE, warning=FALSE, message=FALSE}
# visualizing linear feature importance
library(dotwhisker)
dwplot(mod_1) +
  ggtitle("Linear Model Feature Coefficients") +
  labs(x = "Standardized Effect", y = "Feature Name", subtitle = "Large importance in precipitation, tempurature") +
  theme_classic() +
  geom_vline(xintercept = 0) +
  theme(legend.position = "none")

```


```{r, echo=FALSE, include=FALSE, warning=FALSE, message=FALSE}
# random forest regression
# mod_data
#install.packages("randomForest")
library(randomForest)
library(ggplot2)
rf.fit <- randomForest(num_bikers~tempmax+tempmin+temp+feelslike+dew+humidity+precip+
                         windgust+windspeed+winddir+visibility+solarradiation+severerisk+uvindex,
                       data=mod_data, ntree=1000,
                       keep.forest=FALSE, importance=TRUE)
# rf.fit
rmse <- sqrt(1225608)
rmse

ImpData <- as.data.frame(importance(rf.fit))
ImpData$Var.Names <- row.names(ImpData)
```

```{r, output=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(ImpData, aes(x=Var.Names, y=`%IncMSE`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`%IncMSE`), color="blue") +
  geom_point(aes(size = IncNodePurity), color="dodgerblue4", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="none",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  ggtitle("Random Forest Feature Importances") +
  labs(x = "% Increase in MSE", y = "Variable Name", subtilte = "Precipitation measures prove to be effective predictors of bike traffic")
```

## Killer Plot

```{r, echo = FALSE, output = TRUE, warning=FALSE}
### Killer Plot Stuff

library(grid)


## calculate bearings
library(geosphere)
bike_dta$bearings <- bearing(bike_dta[, c("start_lng", "start_lat")], bike_dta[, c("end_lng", "end_lat")])

#convert to radians
bike_dta$bearings <- -(bike_dta$bearings * pi / 180 - pi/2)
bike_dta$circle_x <- cos(bike_dta$bearings)
bike_dta$circle_y <- sin(bike_dta$bearings)



bike_dta <- subset(bike_dta, start_station_name != "")
bike_dta <- subset(bike_dta, end_station_name != "")


library(tidyverse)
# install.packages('sf')

bike_prepped <- bike_dta %>% group_by(start_station_name, end_station_name) %>% 
  summarise(count = n(),
            bearings = mean(bearings))
bike_prepped$circle_x <- cos(bike_prepped$bearings)
bike_prepped$circle_y <- sin(bike_prepped$bearings)

station_plot_df <- bike_dta %>% group_by(start_station_name) %>% 
  summarise(count = n(),
            start_lat = mean(start_lat),
            start_lng = mean(start_lng))
library(sf)
my_sf <- st_as_sf(station_plot_df, coords = c('start_lng', 'start_lat'))


##Function for scaling counts
opacity_scale <- function(x){(x-min(x))/(max(x)-min(x))}


## Killer Plot Function
killer_plot <- function(df, station_name, station_number = 5){
  
  #subset for desired station
  station_df <- subset(df, start_station_name == station_name)
  
  #find top n stations
  top.n.stations <- station_df[order(-station_df$count),]$end_station_name[1:station_number] 
  
  #set up grid
  grid.newpage()
  vp <- viewport(x = .325, y = 0.65, width = 0.7, height = 0.7)
  pushViewport(vp)
  grid.circle(x = .5, y = .5, r = 0.4, gp = gpar(fill = "black"), default.units = "native")
  grid.circle(x = .5, y = .5, r = 0.38, default.units = "native")
  
  r = 0.38
  
  station_df$circle_x <- station_df$circle_x * r + 0.5
  station_df$circle_y <- station_df$circle_y * r + 0.5
  
  #determine opacity
  station_df$opacity <- opacity_scale(station_df$count)
  
  
  #iterate through trips to draw lines
  for (i in 1:nrow(station_df)){
    angle = station_df[i, "bearings"]
    grid.lines(x = c(.5, station_df$circle_x[i]), y = c(.5, station_df$circle_y[i]), 
               default.units = "native", gp = gpar(lwd = 3, alpha = station_df$opacity[i]))
  }
  
  
  grid.circle(x=.5, y = .5, r = 0.03, gp = gpar(fill = "gray"), default.units = "native")
  
  popViewport()
  vp2 <- viewport(x = .7, y = 0.25, width = 0.6, height = 0.55)
  my_sf$group <- as.factor(ifelse(station_plot_df$start_station_name == station_name, "1", 
                                  ifelse(station_plot_df$start_station_name %in% top.n.stations, "2", "0")))
  #this is not part of the killer plot, so ggplot is okay
  plota <- ggplot(my_sf) + 
    geom_sf(aes(color = group))+
    scale_color_manual(values = c("black", "green", "red"))+
    theme_void()+    
    theme(legend.position = "none")

  print(plota, vp = vp2)
}

## second argument is selecting center station, third is number of stations to highlight in red on the second map.

killer_plot(df = bike_prepped, station_name = names(sort(table(bike_dta$start_station_name), decreasing= TRUE))[8], 15)
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# library(shiny)
# 
# # Define UI for application that draws a histogram
# ui <- fluidPage(
#    
#    # Application title
#    titlePanel("Old Faithful Geyser Data"),
#    
#    # Sidebar with a slider input for number of bins 
#    sidebarLayout(
#       sidebarPanel(
#          sliderInput("highlighted",
#                      "Number of Highlighted Stations:",
#                      min = 1,
#                      max = 20,
#                      value = 5
#           ),
#       
#       selectInput(inputId ="starter_station",
#                   label = "Choose Starting Station",
#                   choices = c(1, 2, 3),
#                   selected = NULL
#       )
#       ),
#       # Show a plot of the generated distribution
#       mainPanel(
#          plotOutput("distPlot")
#       )
#    )
# )
# 
# # Define server logic required to draw a histogram
# server <- function(input, output) {
#   
#   ### Killer Plot Stuff
#   
#   library(grid)
#   
#   # rm(list = ls())
#   
#   
#   ## calculate bearings
# 
#   bike_dta$bearings <- bearing(bike_dta[, c("start_lng", "start_lat")], bike_dta[, c("end_lng", "end_lat")])
#   
#   #convert to radians
#   bike_dta$bearings <- -(bike_dta$bearings * pi / 180 - pi/2)
#   bike_dta$circle_x <- cos(bike_dta$bearings)
#   bike_dta$circle_y <- sin(bike_dta$bearings)
#   
#   
#   
#   bike_dta <- subset(bike_dta, start_station_name != "")
#   bike_dta <- subset(bike_dta, end_station_name != "")
#   
#   bike_prepped <- bike_dta %>% group_by(start_station_name, end_station_name) %>% 
#     summarise(count = n(),
#               bearings = mean(bearings))
#   bike_prepped$circle_x <- cos(bike_prepped$bearings)
#   bike_prepped$circle_y <- sin(bike_prepped$bearings)
#   
#   station_plot_df <- bike_dta %>% group_by(start_station_name) %>% 
#     summarise(count = n(),
#               start_lat = mean(start_lat),
#               start_lng = mean(start_lng))
#   
#   my_sf <- st_as_sf(station_plot_df, coords = c('start_lng', 'start_lat'))
#   
#   
#   ##Function for scaling counts
#   opacity_scale <- function(x){(x-min(x))/(max(x)-min(x))}
#   
#   
#   ## Killer Plot Function
#   killer_plot <- function(df, station_name, station_number = 5){
#     
#     #subset for desired station
#     station_df <- subset(df, start_station_name == station_name)
#     
#     #find top n stations
#     top.n.stations <- station_df[order(-station_df$count),]$end_station_name[1:station_number] 
#     
#     #set up grid
#     grid.newpage()
#     vp <- viewport(x = .325, y = 0.65, width = 0.7, height = 0.7)
#     pushViewport(vp)
#     grid.circle(x = .5, y = .5, r = 0.4, gp = gpar(fill = "black"), default.units = "native")
#     grid.circle(x = .5, y = .5, r = 0.38, default.units = "native")
#     
#     r = 0.38
#     
#     station_df$circle_x <- station_df$circle_x * r + 0.5
#     station_df$circle_y <- station_df$circle_y * r + 0.5
#     
#     #determine opacity
#     station_df$opacity <- opacity_scale(station_df$count)
#     
#     
#     #iterate through trips to draw lines
#     for (i in 1:nrow(station_df)){
#       angle = station_df[i, "bearings"]
#       grid.lines(x = c(.5, station_df$circle_x[i]), y = c(.5, station_df$circle_y[i]), 
#                  default.units = "native", gp = gpar(lwd = 3, alpha = station_df$opacity[i]))
#     }
#     
#     
#     grid.circle(x=.5, y = .5, r = 0.03, gp = gpar(fill = "gray"), default.units = "native")
#     
#     popViewport()
#     vp2 <- viewport(x = .7, y = 0.25, width = 0.6, height = 0.55)
#     my_sf$group <- as.factor(ifelse(station_plot_df$start_station_name == station_name, "1", 
#                                     ifelse(station_plot_df$start_station_name %in% top.n.stations, "2", "0")))
#     #this is not part of the killer plot, so ggplot is okay
#     plota <- ggplot(my_sf) + 
#       geom_sf(aes(color = group))+
#       scale_color_manual(values = c("black", "green", "red"))+
#       theme_void()+    
#       theme(legend.position = "none")
#   
#     print(plota, vp = vp2)
#   }
#    
#    output$distPlot <- renderPlot({
#      killer_plot(df = bike_prepped, station_name = names(sort(table(bike_dta$start_station_name), decreasing= TRUE))[8], input$highlighted)
#    })
# }
# 
# # Run the application 
# shinyApp(ui = ui, server = server)
```

## Conclusion
* Usage of bikes is concentrated heavily in several popular stations
* Generally, bikes are used for short distance and duration trips
* Precipitation is particularly impactful in decreasing the number of bike rides in a day
* Weather data proves to be an effective predictor of bike share usage

